Boosted Regression Trees for Predictive Auto-Tuning

### 2011-12-13 paper/draft1.tex
-sensible +sensitive

% DDC: "sensible to"?  what is meant here? "Sensitive to?"
The empirical approach is very sensible to the choice of instrumented
optimizations \emph{and} to the search method. The size of the search space is
so large that the current best empirical auto-tuners can only consider
highly-specialized functions with a limited set of code transformations and
compiler options, on very few input domains \citep{ganapathi2009case}. Although
searching for good code configurations in this highly-discontinuous space can
be made embarrassingly parallel, it remains a very difficult and prohibitively
expensive combinatorial optimization problem as many variants of the code must
generated, compiled, and benchmarked on specific input domains with meaningful
statistics (that may require multiple runs). Consequently, most proposed
methods prune the space with hard-coded heuristics that offer little
generalization guarantees. This has been the main drawback of the empirical
approach compared to the model-driven approach where good code configurations
can be directly derived from the analytical model.

@@

In this paper, we propose to \emph{learn the model}, empirically, (XXX:
we'll describe what this means in more details later in the paper, riiiighhht
???) instead of deriving it analytically. By learning the model, one can
hope to achieve elements of the best of both approaches and capitalize on the large body of
literature from the machine learning community.

Although various statistical prediction techniques have been applied with
success at compile-time for general programs on various CPU architectures
\citep{monsifrot2002machine, stephenson2003meta, yotov2003comparison,
kulkarni2004fast, cooper2005acme, franke2005probabilistic,
hutter2006performance, cavazos2007rapidly, cavazos2008intelligent,
hartono2009annotation, park2011evaluation, fursin2008milepost}, there is, to
our knowledge, no fast predictive auto-tuning work that has tried to cover
simultaneously: (a) more recent multi-core architectures like GPUs
\citep{schaa2009exploring}, (b) high-performance domain-specific libraries
\citep{nukada2009auto, li2009note, kamil2010auto}, (c) at run-time
\citep{klockner2011pycuda, pinto+cox:2011gcg}, and (d) on \emph{many} input
domains \citep{liu2009cross, grauer2011optimizing}.

XXX WIIFY here ??? (what's in their for you) XXX in text, say what's in this paper for:
- compiler people
- libs people
- ml people
- tools people
- cv people

@@

% XXX: our contributions
% XXX: WIIFY
% XXX: organisation of the paper

@@

\subsection{Case Study}
Current languages and software engineering practices have conspired to make it difficult
to provide high-performance libraries for general mathematical operations on GPU devices.
An important approach in implementing high-performance GPU codes has been auto-tuning,
which is fundamentally slow. Auto-tuning requires running a large number of
sub-compilations and representative computations to {\em measure} which
of several possible implementations is fastest. General-purpose libraries such
as CUBLAS and CUFFT XXX hide this overhead by auto-tuning offline and using
intimate knowledge of the mathematical problem being solved as well as the
kernel implementations to heuristically choose a good computation strategy quickly when a
user asks for e.g. a matrix product of particular arguments.
Designing and testing libraries of this form is time-consuming, difficult, and provides little
reusable infrastructure to assist in the implementation of different computations.
For instance, even if one has already developed an auto-tuned library for dense matrix algebra, 
one must basically start from scratch to develop an auto-tuned library for
stencil computations or sparse matrix operations.

@@

Our contribution of the present work is to show that boosted regression tree models,
a powerful machine learning technique for function approximation, can meet the
requirements of this application.
We call this approach {\em predictive} auto-tuning to contrast it with the standard measurement-based approach which we will call {\em empirical} auto-tuning.
XXX: We show that the kernel from the GCG chapter can be modeled on a variety of hardware and get good performance compared to empirical auto-tuning.

\cite{volkov+demmel:2008}
\cite{vuduc:2003} % has some stuff about ML for autotuning in ch. 9

>>>>

The empirical approach is very sensitive to the choice of instrumented
optimizations \emph{and} to the search method. The size of the search space is
so large that the current best empirical auto-tuners can only consider
highly-specialized functions with a limited set of code transformations and
compiler options, on very few input domains \citep{ganapathi2009case}. Although
searching for good code configurations in this highly-discontinuous space can
be made embarrassingly parallel, it remains a very difficult and prohibitively
expensive combinatorial optimization problem as many variants of the code must
generated, compiled, and benchmarked on specific input domains with meaningful
statistics (that may require multiple runs). Consequently, most proposed
methods prune the space with hard-coded heuristics that offer little
generalization guarantees. This has been the main drawback of the empirical
approach compared to the model-driven approach where good code configurations
can be directly derived from the analytical model.

@@

In this paper, we propose to \emph{learn the model} using non-linear regression
modelling techniques instead of constructing a model manually. By learning the model, one can
hope to achieve elements of the best of both approaches: XXX.
%and capitalize on the large body of
%literature from the machine learning community.

Various statistical prediction techniques have been applied with
success at compile-time for general programs on various CPU architectures
\citep{monsifrot2002machine, stephenson2003meta, yotov2003comparison,
kulkarni2004fast, cooper2005acme, franke2005probabilistic,
hutter2006performance, cavazos2007rapidly, cavazos2008intelligent,
hartono2009annotation, park2011evaluation, fursin2008milepost}.
Relative to this work our contribution is to show
how to do fast predictive auto-tuning that satisifies the requirements to:
(a) handle the variety of recent multi-core architectures like GPUs \citep{schaa2009exploring},
(b) provide high-performance domain-specific libraries \citep{nukada2009auto, li2009note, kamil2010auto},
(c) that select good implementations at run-time \citep{klockner2011pycuda, pinto+cox:2011gcg}, and
(d) for the full input domain of a library routine \citep{liu2009cross, grauer2011optimizing}.
% XXX: are these last references appropriate
% XXX: there are 4 versions of the stencil work - consolidate to just the full journal version

\iffalse
XXX WIIFY here ??? (what's in their for you) XXX in text, say what's in this paper for:
- compiler people
- libs people
- ml people
- tools people
- cv people
\fi

% XXX: organisation of the paper?

@@

@@

\subsection{Case Study}
Current languages and software engineering practices have conspired to make it difficult
to provide high-performance libraries for general mathematical operations on GPU devices.
An important approach in implementing high-performance GPU codes has been auto-tuning,
which is fundamentally slow. Auto-tuning requires running a large number of
sub-compilations and representative computations to {\em measure} which
of several possible implementations is fastest. General-purpose libraries such
as CUBLAS and CUFFT hide this overhead by auto-tuning offline and using
intimate knowledge of the mathematical problem being solved as well as the
kernel implementations to heuristically choose a good computation strategy quickly when a
user asks for e.g. a matrix product of particular arguments.
Designing and testing libraries of this form is time-consuming, difficult, and provides little
reusable infrastructure to assist in the implementation of different computations.
For instance, even if one has already developed an auto-tuned library for dense matrix algebra, 
one must basically start from scratch to develop an auto-tuned library for
stencil computations or sparse matrix operations.

@@

Our contribution of the present work is to show that boosted regression tree models,
a powerful machine learning technique for function approximation, can meet the
requirements of this application.
We call this approach {\em predictive} auto-tuning to contrast it with the standard measurement-based approach which we will call {\em empirical} auto-tuning.
We show that the kernel from the GCG chapter can be modeled on a variety of hardware and get good performance compared to empirical auto-tuning.
%XXX: this has no flow
